\Large \bf A first look at auditing in a blockchain world

### 2018-09-25 auditing/Latex/sections/02-body.tex
Audit clean up

% !TEX root = ../main.tex
\section{Key Issue: Existence of Assets and Occurence of Transactions}

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Finality.} While blockchains are touted as immutable, the finality property (like all security properties) is subject to assumptions. Immutability of a blockchain is subject to consensus taken across miners according to computational ability. Consensus is not instant: a transaction might be included and then quickly dropped as consensus forms between different proposed chains. And it is never guaranteed to be final: an agreement within the computational majority of miners can unroll past transactions, For example, Ethereum{'}s miners branched the main blockchain to modify some past transactions to reverse a hacked DApp transactions~\cite{dupont2017experiments}. The challenge of finality was an important design consideration in the Bank of Canada{'}s Project Jasper~\cite{chapman2017project} which sought to establish a Distributed Wholesale Payment Systems. ``Project Jasper was structured so that a transfer of [digital currency] was equivalent to a full and irrevocable transfer of the underlying claim on central bank deposits.'' Given that on the Ethereum blockchain, transactions could be reversed, the work around for Project Jasper was to implement a ``design feature (which) relates to the issuance of (digital currency) and is therefore independent of the platforms upon which Jasper was built.''.

The issue of revocability is not uncommon for accountants. In many cases, sales agreements provide customers with the option to return merchandise within a pre-established period. When recording revenue, the accountant much estimate the percentage of returns and factor this into the amount of revenue to be recorded. The same concept could be applied for the issue of finality. The firm could estimate the amount of returns or reversals that are likely to occur and factor this into their transaction recognition. Alternatively, auditors could establish a generally accepted threshold (for instance, number of days) after which a transaction would be considered final.  %we could discuss the practice of transaction confirmation for 6+ blocks here as well

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Transaction Pointers.} Blockchain transactions need to be uniquely identified to be pointed at by secondary financial records. Due to an implementation fault in Bitcoin, transactions broadcasted with one identifier might end up in the blockchain with a different identifier -- Transaction Malleability --, which could lead a firm to conclude a transaction did not take place when in fact it did, but under an unexpected identifier~\cite{andrychowicz2015malleability}. Some firms went bankrupt when their automated system kept honouring refund requests from a malicious entity claiming the refunds were not going through~\cite{trautman2014virtual,decker2014bitcoin}. The larger lesson here is that auditors cannot always safely abstract away low level implementation details.

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Occurrence.} \todo{This contradicts some of the points mentioned in Transaction Pointers and Finality, also the point made here is also mentioned there. My suggestion is to remove this Issue or merge it with one of the above} In order to validate that cryptoasset transactions actually took place during the year, auditors will look to reproduce the transactions carried out. By obtaining a listing of the client{'}s cryptographic keys, the blockchain can provide a complete history of the transactions that occurred during the year. Due to the immutable nature of the blockchain, the auditor can be assured that the transaction did take place as described. However, the auditor will still need to validate meta-information surrounding the transaction (see above) to validate the transaction{'}s reasonability.

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway:} This section demonstrates that due to the immutable nature of the blockchain and its ability to report the totality of transactions conducted during the period, this technology provides a record upon which the auditor can obtain evidence to ascertain the occurrence of transactions. However, the auditor will continue to need to rely on external sources such as a verbose secondary ledger to validate the legitimacy and business purpose of those transactions.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Cryptographic keys.} For most cryptoassets, the asset is considered owned by Alice if Alice possesses a private signing key that can be used to sign a transfer of the asset. Through decentralized apps, alternative notions of ownership are possible to define, but this idea of a signing key is foundational and seen in bitcoin, ether, ERC20 tokens, etc. (not to mention earlier e-cash proposals dating back to the 1980s). Thus demonstrating knowledge of this key is necessary (but not sufficient, as we will discuss shortly) to demonstrating ownership. The most direct technique is to use a zero knowledge proof of knowledge of this private key, and staple in some information identifying the context of the proof. For standard proofs, this is cryptographically equivalent to simply signing a challenge message with the key\footnote{A Schnorr Sigma-protocol with the challenge hashed in using Fiat-Shamir is exactly a Schnorr signature and closely related to an ECDSA signature.\todo{CITE}}. Folklore protocols of sending small cash amounts from an allegedly owned account to the auditor to demonstrate control are also commonly noted in the literature (and used in at least one occasion \todo{[cite]}). This offers similar security but adds ethical complexities for the auditor.
\todo{ It requires more [cryptography] citations here! ~\cite{swire_uses_nodate,blaze1998compliance,kohlas_reasoning_nodate} }
% What to cite for the [cite] placeholder
We note that while this cryptographic proof is necessary, it is not sufficient. What it proves is that the purported owner has access to the person holding the signing key. A malicious company might arrange for the owner of cryptoassets to engage in signing statements or moving test amounts fraudulently on their behalf. This issue is not new: an insolvent retail store might borrow inventory from elsewhere to inflate its assets during an audit, which will, in this case, involve a physical visit and inventory check by the auditor. Auditors mitigate this by arranging a common date for all audits of physical inventory, and similarly, cryptographic audits might be synchronized on a fixed schedule to prevent the same assets from being counted for different companies in different audits~\cite{dagher2015provisions}.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Self-custodianship.} Cryptocurrency advocates point to its non-reliance on trusted third parties as the key to its appeal \todo{[cite: CHI paper]}. Thus, the use of a custodian for cryptoassets (or more specifically, the private keys controlling the assets) is controversial. The advantage of a custodian is that it might more quickly specialize in security than end users, and this is certainly true for some users \todo{[cite: MIT]}. Self-custodianship of non-digital assets, such as diamonds for a jewelry retailer or cash for a currency exchange, is already a concern for financial auditors. Further, custodianship over cryptographic keys is a financial factor in other sectors, such as certificate authorities like Verisign or Symantec which maintain keys critical to HTTPS and DNSSEC \todo{[cite]}. To date, no blockchain custodian or exchange has been able to produce a report that supports the reliability of their internal controls in order to provide auditors with comfort over the sufficiency of their systems. Therefore, auditors cannot rely on the internal controls present at custodians to obtain comfort over the ownership assertion.

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway:} In order for auditors to validate ownership, they must rely on cryptographic proofs as a first step. However, in order to avoid double-counting of keys, an industry standard common date should be arranged to provide a generally agreed upon ``state of the world'' where keyholders can demonstrate ownership.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Geographical Variation.} Because cryptoassets(\eg bitcoin) can be moved digitally without any geographic restrictions, it should be the case that any variation in cryptoasset prices across exchanges would be consumed by arbitrageurs. However, empirically this is not the case due to fees, settlement delays, and other frictions with exchange services \todo{[cite]}.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway:} While the valuation of cryptoassets with large trading volumes like Bitcoin and Ether would be fairly straightforward by referring to values in an active market, the valuation of coins with low liquidity may be more difficult. However, by using financial modelling using different assumptions for volatility, this challenge is not insurmountable.

@@

\paragraph{Issue: Reliability of Blockchain Records.} Another challenge for auditors is determining what type of evidence or records constitute sufficient and appropriate audit evidence. In other words, an auditor must be satisfied that the records on the blockchain are reliable. How can we satisfy auditors that blockchain records are reliable
 \todo{This subsection needs some cutting, has duplicate points and not obvious what the main objective is. there are some comments on the google docs regarding this subsection that should be addressed}

\begin{enumerate}
\item[$\bullet$] How can we rely on blockchain? Has anyone tested it? For example, maybe Bitcoin is reliable due to large user base but what about more niche blockchain? How can I validate accuracy of mining procedures? Reliability of transaction records? How do I know that changes to blockchain code made and not undermine reliability?
\end{enumerate}

Traditionally, auditors depend on obtaining audit evidence from external parties as it is deemed as more objective or trustworthy than internal firm documents. For example, auditors will obtain confirmation from a bank of the amount of cash held at year end (a third party) rather than rely on internal records of banking transactions. Auditors traditionally rely on banks, investment companies, and lawyers as conventional sources of external confirmations.  However, a risk always exists that there is fraud or error at these institutions.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Financial Literacy of Technology Professionals} While auditors must learn more about the technology, managers of companies in the blockchain space must also learn about the internal controls required to safeguard against errors, misappropriation of assets and fraud. Given that financial statements are ultimately the responsibility of management and that senior management must certify the internal controls as part of the financial statement disclosures, managers have an incentive to implement a control environment that promotes a strong tone at the top, ethical conduct, and oversight of the financial reporting function. %TODO: could say something about/cite : blaze1998compliance

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Internal Controls.} It is necessary for a firm to be able to demonstrate the existence and ownership of its crypto-assets, however it must further demonstrate that it has adequate procedures in place to prevent or detect fraud and theft. Consider a firm holding DAO tokens. Internally, it must ensure proper controls over the signing keys these DAO tokens are assigned to. External to the firm, the DAO application maintaining these tokens on Ethereum{'}s blockchain must itself be secure (which in the case of the DAO, it was not~\cite{dupont2017experiments}). Auditing internal procedures over cryptographic keys is not unprecedented---maintaining certificate authority keys is critical to some security firm{'}s financial prospects (a security breach at one firm, \textit{DigiNotar} led to its bankruptcy ~\cite{zetter2011diginotar}). Symantec{'}s self-custodianship over these keys involves elaborate security ceremonies whenever they are required \todo{[Cite]}.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Detection of Fraud.} An auditor conducting an audit in accordance with GAAS is responsible for obtaining reasonable assurance that the financial statements taken as a whole are free from material misstatement, whether caused by fraud or error. Owing to the inherent limitations of an audit, there is an unavoidable risk that some material misstatements or fraud may not be detected, even though the audit is properly planned and performed in accordance with GAAS. Due to the complex and rapidly changing nature of this sector, auditors are especially weary of fraud risk in this area. 
However, not all frauds that occur this space are new kinds of fraud. For instance, Ponzi schemes have existed since they were perfected by Charles Ponzi in the early twentieth century~\cite{ponzi2001rise}. In 2017, OneCoin was found to have raised over \$350M of funds from investors for an ICO that was a Ponzi scheme~\cite{atlanticCryptoPonzi2017}. Stories like this~\cite{vasek2015analyzing} only reinforce the perception in the auditing community of the dangers of operating in this space. However, auditors can protect themselves by ensuring that they only work with reputable clients. Managers can demonstrate their commitment to sound business practices by adopting controls that ensure KYC and AML rules are adhered to.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
\section{Discussion and Research Agenda}
\todo{Check out the notes on the google docs regarding this section}

@@

\todo{Check out the notes on the google docs regarding this section}

>>>>

% !TEX root = ../main.tex
\section{Key Issue: Existence of Assets}

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Finality.} While blockchains are touted as immutable, the finality property (like all security properties) is subject to assumptions. Immutability of a blockchain is subject to consensus taken across miners according to computational ability. Consensus is not instant: a transaction might be included and then quickly dropped as consensus forms between different proposed chains. And it is never guaranteed to be final: an agreement within the computational majority of miners can unroll past transactions, For example, Ethereum{'}s miners branched the main blockchain to modify some past transactions to reverse a hacked DApp transactions~\cite{dupont2017experiments}. The challenge of finality was an important design consideration in the Bank of Canada{'}s Project Jasper~\cite{chapman2017project} which sought to establish a Distributed Wholesale Payment Systems. ``Project Jasper was structured so that a transfer of [digital currency] was equivalent to a full and irrevocable transfer of the underlying claim on central bank deposits.'' Given that on the Ethereum blockchain, transactions could be reversed, the work around for Project Jasper was to implement a ``design feature (which) relates to the issuance of (digital currency) and is therefore independent of the platforms upon which Jasper was built.''

The issue of revocability is not uncommon for accountants. In many cases, sales agreements provide customers with the option to return merchandise within a pre-established period. When recording revenue, the accountant much estimate the percentage of returns and factor this into the amount of revenue to be recorded. The same concept could be applied for the issue of finality. The firm could estimate the amount of returns or reversals that are likely to occur and factor this into their transaction recognition. Alternatively, auditors could establish a generally accepted threshold (for instance, 6 blocks is commonly used in bitcoin) after which a transaction would be considered final.  %we could discuss the practice of transaction confirmation for 6+ blocks here as well

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Transaction Pointers.} Blockchain transactions need to be uniquely identified to be pointed at by secondary financial records. Due to an implementation fault in Bitcoin, transactions broadcast with one identifier might end up in the blockchain with a different identifier. Known as transaction malleability, firms might erroneously conclude a transaction did not take place when in fact it did under an unexpected identifier~\cite{andrychowicz2015malleability}. Some firms went bankrupt when their automated system kept honouring refund requests from a malicious entity claiming the refunds were not going through~\cite{trautman2014virtual,decker2014bitcoin}. The larger lesson here is that auditors cannot always safely abstract away low-level implementation details.

%---------+++++++++++++++++++-----------------
%\paragraph{Issue: Occurrence.} \todo{This contradicts some of the points mentioned in Transaction Pointers and Finality, also the point made here is also mentioned there. My suggestion is to remove this Issue or merge it with one of the above} In order to validate that cryptoasset transactions actually took place during the year, auditors will look to reproduce the transactions carried out. By obtaining a listing of the client{'}s cryptographic keys, the blockchain can provide a complete history of the transactions that occurred during the year. Due to the immutable nature of the blockchain, the auditor can be assured that the transaction did take place as described. However, the auditor will still need to validate meta-information surrounding the transaction (see above) to validate the transaction{'}s reasonability.

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway.} This section demonstrates that due to the immutable nature of the blockchain and its ability to report the totality of transactions conducted during the period, this technology provides a record upon which the auditor can obtain evidence to ascertain the occurrence of transactions. However, the auditor will continue to need to rely on external sources such as a verbose secondary ledger to validate the legitimacy and business purpose of those transactions.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Cryptographic keys.} For most cryptoassets, the asset is considered owned by Alice if Alice possesses a private signing key that can be used to sign a transfer of the asset~\cite{Swi97,BFS98,KM99}. Through decentralized apps, alternative notions of ownership are possible to define, but this idea of a signing key is foundational and seen in bitcoin, ether, ERC20 tokens, etc. (not to mention earlier e-cash proposals dating back to the 1980s). Thus demonstrating knowledge of this key is necessary (but not sufficient, as we will discuss shortly) to demonstrating ownership. The most direct technique is to use a zero knowledge proof of knowledge of this private key, and staple in some information identifying the context of the proof. For standard proofs, this is cryptographically equivalent to simply signing a challenge message with the key\footnote{A Schnorr Sigma-protocol with the challenge hashed in using Fiat-Shamir is exactly a Schnorr signature and closely related to an ECDSA signature.}. Folklore protocols of sending small cash amounts from an allegedly owned account to the auditor to demonstrate control are also commonly noted in the literature. This offers similar security but adds ethical complexities for the auditor.

We note that while this cryptographic proof is necessary, it is not sufficient. What it proves is that the purported owner has access to the person holding the signing key. A malicious company might arrange for the owner of cryptoassets to engage in signing statements or moving test amounts fraudulently on their behalf. This issue is not new: an insolvent retail store might borrow inventory from elsewhere to inflate its assets during an audit, which will, in this case, involve a physical visit and inventory check by the auditor. Auditors mitigate this by arranging a common date for all audits of physical inventory, and similarly, cryptographic audits might be synchronized on a fixed schedule to prevent the same assets from being counted for different companies in different audits~\cite{dagher2015provisions}.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Self-custodianship.} Cryptocurrency advocates point to its non-reliance on trusted third parties as the key to its appeal~\cite{GCL16}. Thus, the use of a custodian for cryptoassets (or more specifically, the private keys controlling the assets) is controversial. The advantage of a custodian is that one firm can specialize in security rather than all end users. Self-custodianship of non-digital assets, such as diamonds for a jewelry retailer or cash for a currency exchange, is already a concern for financial auditors. Custodianship over cryptographic keys is a factor in other sectors, such as certificate authorities like Verisign or Symantec which maintain keys critical to HTTPS and DNSSEC~\cite{clark2013sok}. To date, no blockchain custodian or exchange has been able to produce a report that supports the reliability of their internal controls in order to provide auditors with comfort over the sufficiency of their systems. Therefore, auditors cannot rely on the internal controls present at custodians to obtain comfort over the ownership assertion.

% Might have to soften "to date..." sentence

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway.} In order for auditors to validate ownership, they must rely on cryptographic proofs as a first step. However, in order to avoid double-counting of keys, an industry standard common date should be arranged to provide a generally agreed upon ``state of the world'' where keyholders can demonstrate ownership.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Geographical Variation.} Because cryptoassets(\eg bitcoin) can be moved digitally without any geographic restrictions, it should be the case that any variation in cryptoasset prices across exchanges would be consumed by arbitrageurs. However, empirically this is not the case due to fees, settlement delays, and other frictions with exchange services~\cite{KS17}. Reporting standards do not account for geographic variation.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Takeaway.} While the valuation of cryptoassets with large trading volumes like Bitcoin and Ether would be fairly straightforward by referring to values in an active market, the valuation of coins with low liquidity may be more difficult. However, by using financial modelling using different assumptions for volatility, this challenge is not insurmountable.

@@

%---------+++++++++++++++++++-----------------

%\paragraph{Issue: Reliability of Blockchain Records.} Another challenge for auditors is determining what type of evidence or records constitute sufficient and appropriate audit evidence. In other words, an auditor must be satisfied that the records on the blockchain are reliable. How can we satisfy auditors that blockchain records are reliable
% \todo{This subsection needs some cutting, has duplicate points and not obvious what the main objective is. there are some comments on the google docs regarding this subsection that should be addressed}
%
%\begin{enumerate}
%\item[$\bullet$] How can we rely on blockchain? Has anyone tested it? For example, maybe Bitcoin is reliable due to large user base but what about more niche blockchain? How can I validate accuracy of mining procedures? Reliability of transaction records? How do I know that changes to blockchain code made and not undermine reliability?
%\end{enumerate}
%
%Traditionally, auditors depend on obtaining audit evidence from external parties as it is deemed as more objective or trustworthy than internal firm documents. For example, auditors will obtain confirmation from a bank of the amount of cash held at year end (a third party) rather than rely on internal records of banking transactions. Auditors traditionally rely on banks, investment companies, and lawyers as conventional sources of external confirmations.  However, a risk always exists that there is fraud or error at these institutions.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Financial Literacy of Technology Professionals.} While auditors must learn more about the technology, managers of companies in the blockchain space must also learn about the internal controls required to safeguard against errors, misappropriation of assets and fraud. Given that financial statements are ultimately the responsibility of management and that senior management must certify the internal controls as part of the financial statement disclosures, managers have an incentive to implement a control environment that promotes a strong tone at the top, ethical conduct, and oversight of the financial reporting function. %TODO: could say something about/cite : blaze1998compliance

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Internal Controls.} It is necessary for a firm to be able to demonstrate the existence and ownership of its crypto-assets, however it must further demonstrate that it has adequate procedures in place to prevent or detect fraud and theft. Consider a firm holding DAO tokens. Internally, it must ensure proper controls over the signing keys these DAO tokens are assigned to. External to the firm, the DAO application maintaining these tokens on Ethereum{'}s blockchain must itself be secure (which in the case of the DAO, it was not~\cite{dupont2017experiments}). Auditing internal procedures over cryptographic keys is not unprecedented---maintaining certificate authority keys is critical to some security firm{'}s financial prospects (a security breach at one firm, \textit{DigiNotar} led to its bankruptcy ~\cite{zetter2011diginotar}). Symantec{'}s self-custodianship over these keys involves elaborate security ceremonies whenever they are required~\cite{Goo12}.

@@

%---------+++++++++++++++++++-----------------
\paragraph{Issue: Detection of Fraud.} An auditor conducting an audit in accordance with GAAS is responsible for obtaining reasonable assurance that the financial statements taken as a whole are free from material misstatement, whether caused by fraud or error. Owing to the inherent limitations of an audit, there is an unavoidable risk that some material misstatements or fraud may not be detected, even though the audit is properly planned and performed in accordance with GAAS. Due to the complex and rapidly changing nature of this sector, auditors are especially weary of fraud risk in this area. 
Not all frauds that occur this space are new kinds of fraud. For instance, Ponzi schemes have existed since they were perfected by Charles Ponzi in the early twentieth century~\cite{ponzi2001rise}. In 2017, OneCoin was found to have raised over \$350M of funds from investors for an ICO that was a Ponzi scheme~\cite{atlanticCryptoPonzi2017}. Stories like this~\cite{vasek2015analyzing} only reinforce the perception in the auditing community of the dangers of operating in this space. However, auditors can protect themselves by ensuring that they only work with reputable clients. Managers can demonstrate their commitment to sound business practices by adopting controls that ensure KYC and AML rules are adhered to.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
\section{Discussion and Research Agenda}

%\todo{Check out the notes on the google docs regarding this section}

@@

Blockchain technology provides novel, non-trivial issues for financial auditors, however these issues are in no way insurmountable and in many ways parallel challenging conditions seen in other sectors. We conclude that many entities in this space are auditable. We also wish to note that the firm is ultimately the entity seeking the audit, auditors are not seeking the firm, and therefore the onus to ensure financial disclosures are readably auditable by non-specialists auditors really lies with the firm itself. For example, a firm might not want to use a custodian for cryptocurrencies for philosophical reasons, and overtime, the technical literacy of auditors will advance to where self-custodianship may be commonplace, however firms seeking audits in the short-term may need to invest time and resources into internal controls, financial reporting mechanisms, and third parties to enable less friction for auditors.
